---
title: "Diet analysis of herring predation in Puget Sound"
output: html_notebook
---


Load libraries
```{r load libraries, include=FALSE}
source("libraries.R")
source("test_herring.R")
source("test_herring_raw_data.R")
source("clean_herringdata.R")
source("clean_herringpreddata.R")
source("make_map.R")
source("get_basin_coord.R")
source("get_raw_data.R")
source("gam_table.R")
```


Check full diet data set for herring prey and move to herring diet folder, add full reference

```{r check data, include=FALSE}

data.files <- list.files(path="~/herringdiet/all_diet_data",pattern = "*.xlsx", full.names = TRUE)

#Moves data files to herring data
lapply(data.files,test_herring)

herring.predators <- read_csv("herring_predators.csv") %>% 
  pull(atlantis_pred_group)

#test for data files that have herring predators but do not report herring
lapply(data.files,test_herring_preds, herring.predators)


data.files.raw <- list.files(path="~/herringdiet/raw_diet_data",pattern = "*.csv", full.names = TRUE)

#Moves data files to herring data
lapply(data.files.raw,test_herring_raw_data)


```


Include raw herring data

```{r}
# add raw data clean files

#source("Kwiath_diets.R")
#source("clean_raw_data.R")

raw.data.files <- list.files("~/herringdiet/raw_diet_data_rev", full.names = TRUE)

raw.data <- lapply(raw.data.files,get_raw_data) %>% 
  bind_rows %>% 
   dplyr::select(-loc_date_index, -num_fish, -taxa_predator, -loc_index) %>% 
  rename(predator_taxa = predator)

raw.data %>% 
  write_csv("~/herringdiet/herring_raw_diet_data.csv")
```



```{r clean herring data}

herring.pred.files <- list.files("~/herringdiet/herring_pred_data_rev", full.names = TRUE)

herring.pred.data <- lapply(herring.pred.files,clean_herringpreddata) %>% 
  bind_rows %>% 
  dplyr::select(-data) %>% 
  mutate(variable = gsub("ocurrence", "occurrence", variable))


raw.data  <- fread("herring_raw_diet_data.csv") %>% 
  as_tibble

data.folder <- "~/herringdiet/herring_data_rev"

data.files <- list.files(path=data.folder,pattern = "*.xlsx", full.names = TRUE)

herring.clean.data <- lapply(data.files, clean_herringdata, diet.names="diet_file_list.csv", predator.list = "herring_predator_names.csv") %>% 
  bind_rows %>% 
  bind_rows(herring.pred.data)

all.herring.data <- herring.clean.data %>% 
   dplyr::select(-file_name, -full_file_name, -data, -index, -value, -predator_name) %>% 
  bind_rows(raw.data)

diet.file.names <- read_csv("diet_file_list.csv") %>% 
  rename(file_name = FILE, reference = CITA)

herring.clean.data %>% 
  mutate(file_name=gsub(".xlsx","",file_name)) %>% 
  left_join(diet.file.names, by="file_name") %>% 
  distinct(file_name, reference) %>% 
  arrange(reference) %>% 
  #filter(is.na(reference)) %>% 
  write_csv("herring_references.csv")

  #using all predators
  #selected groups are 
  #pred.groups <- c("HSL","PIN","SAL","SB","FMM","CI","CO","CU","DOG","PIS","ROC","MRO","SMD","SP")
  
  
all.herring.data %>% 
  #filter(is.na(latitude) & !is.na(longitude)) %>% 
  distinct(location, latitude,longitude) %>% 
  arrange(location) %>% 
  write_csv("herring_locations.csv")

all.herring.data %>% 
  #filter(is.na(latitude) & !is.na(longitude)) %>% 
  distinct(atlantis_pred_group) %>% 
  arrange(atlantis_pred_group) %>% 
  write_csv("herring_predators.csv")

#join with Raw data and convert predator_size based on units


clean.all.herring.data <- all.herring.data %>% 
  mutate(predator_stage= tolower(predator_stage)) %>% 
   mutate(predator_stage = if_else(predator_stage=="juvenileandadult","adult",
                                   if_else(predator_stage=="adults","adult", predator_stage))) %>% 
 mutate(predator_stage = if_else(is.na(predator_stage),"adult",predator_stage)) %>%
  mutate(atlantis_pred_group = if_else(atlantis_pred_group == "STEELHEAD", "ST", atlantis_pred_group)) %>% 
   mutate(atlantis_pred_group = if_else(atlantis_pred_group == "MYCTOPHID", "MYC", atlantis_pred_group)) %>% 
  mutate(atlantis_pred_group=if_else(atlantis_pred_group %in% c("ROCKFISH","DVR","MRO"),"ROC",atlantis_pred_group)) %>% 
  mutate(predator_size=if_else(size_units=="in",predator_size*25.4,
                               if_else(size_units=="cm",predator_size*10,
                                       if_else(size_units=="lbs",predator_size*453.592,predator_size)))) %>% 
  mutate(size_units=if_else(size_units=="lbs","g",
                            if_else(size_units=="cm","mm",
                                    if_else(size_units=="in","mm", size_units)))) %>% 
  mutate(size_units = if_else(size_units=="NA",NA_character_,size_units)) %>% 
  mutate(predator_size = if_else(atlantis_pred_group=="HSL" & size_units=="g",(((predator_size/1000)*5095)^(1/2.550)),predator_size)) %>% 
  mutate(size_units = if_else(atlantis_pred_group=="HSL" & size_units=="g","mm",size_units)) #this is data for predators that didn't have herring reported as prey
 # filter(proportion>0)
#used relationship for HSL lenght from Boulva & McLaren 1979 , W(kg) = L(cm)^2.550/5695

clean.all.herring.data %>% 
  filter(!atlantis_pred_group%in% c("BIRD","BD","FRESHWATER","LAMPREY","CSL","HAP")) %>% 
  filter(year>=1970) %>% 
  write_csv("herring_diet_data.csv")



```


Assign each data point to a basin

```{r assign basin, echo=TRUE}

herring.locations <-"herring_diet_data.csv"
boundary.file <- "shapefile_4071.shp"


herring.basin.data <- get_basin_coord(herring.locations, boundary.file) %>% 
  mutate(sample_source = if_else(sample_source=="scat","scats", sample_source),
         variable = if_else(variable=="freq_ocurrence","freq_occurrence",variable))

write_csv(herring.basin.data, "herring_diet_basin_data.csv")

```


Plot sampling points and map

```{r make map, echo=FALSE}

 #https://boundingbox.klokantech.com/
#-123.7904,47.0152,-121.9301,49.2511 Puget Sound

#shape.file <- "USMaritimeLimitsNBoundaries.shp"
boundary.file <- "shapefile_4071.shp"
file.name <- "herring_map.png"
scale.factor <- 2 #map scale
bar.position  <- "tl" #position scale bar
herring.locations <-"herring_diet_basin_data.csv"

  target.crs <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  
 #read Puget Sound boundary polygons
  
  print("read Puget Sound boundary polygons")
  boundary.shape <- st_read(paste0("~/herringdiet/shapefiles/",boundary.file))
  boundary.shape.crs <- st_transform(boundary.shape, target.crs)
  
  #read world map
  print("read world map")
  world <- ne_countries(scale = "large", returnclass = "sf")
  class(world)
 
  #read point data
  print("read point data")
  
  herring.data <- read_csv(herring.locations) %>% 
    distinct(latitude,longitude)
  
  coordinates(herring.data) <- c("longitude","latitude")
  
  proj4string(herring.data)  <- target.crs

  herring.data.sf <- st_as_sf(herring.data)
  
  #eliminate data points outside Puget Sound basins
  print("eliminate data points outside Puget Sound basins")
  coord.intersect <- raster::intersect(herring.data, boundary.shape.crs)
  
  coord.intersect.sf <- st_as_sf(coord.intersect)
  
  print("set colors")
  col.pal <- c(redmonder.pal(8,"qMSOMed"),redmonder.pal(8,"qMSOPap")[8])
  
  print("generate map")

  max.long <- -122
  min.long <- -125
  min.lat <- 47.1
  max.lat <- 49.7
  
  model.map <- ggplot(data=world) +
    geom_sf()+
    geom_sf(data = boundary.shape.crs, aes(fill=NAME)) +
    geom_sf(data = coord.intersect.sf, alpha = .75, colour = "black", size = 0.5) +
    coord_sf(xlim = c(min.long, max.long), ylim = c(min.lat, max.lat), expand = TRUE) +
    scale_fill_manual(values = col.pal, name="Sub-basin") + 
    annotation_north_arrow(location = "tr", which_north = "true", 
                           pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"),
                           style = north_arrow_fancy_orienteering) +
    annotation_scale(location = "tr", width_hint = 0.2,
                     pad_x = unit(0.1, "cm"), pad_y = unit(0.25, "cm")) +
  #  annotate(geom = "text", x = -122.15, y = 47.6, label = "Seattle", 
  #  fontface = "bold", color = "darkblue", size = 4) +
    xlab("Lon")+
    ylab("Lat")+
    theme_minimal()
  
  #inset map
  #https://geocompr.github.io/post/2019/ggplot2-inset-maps/
  
  data("us_states", package = "spData")
  
  #extract bounding box
  
  point.bb  <- st_as_sfc(st_bbox(coord.intersect.sf))
  
  max.long <- -74
  min.long <- -126
  min.lat <- 28
  max.lat <- 52

  inset.map <- ggplot(data=world) +
    geom_sf() +
    geom_sf(data = point.bb, fill = NA, color = "darkred", size = 1) +
    coord_sf(xlim = c(min.long, max.long), ylim = c(min.lat, max.lat), expand = TRUE) +
    annotate(geom = "text", x = -100, y = 51, label = "Canada", 
    fontface = "plain", color = "darkblue", size = 5) +
    annotate(geom = "text", x = -100, y = 45, label = "USA", 
    fontface = "plain", color = "darkblue", size = 5) +
    theme_no_axes()
  
  model.inset.map <- ggdraw() +
  draw_plot(model.map) +
  draw_plot(inset.map, x = 0.1, y = 0.02, width = 0.3, height = 0.3)

  #save_plot(file.name, model.inset.map, ncol=1, base_width = 7, base_height = 4)

  ggsave(file.name, model.inset.map, device="png",width=10,height=10, dpi=400)

#make_map(boundary.file, file.name, scale.factor, bar.position, max.long, min.long, min.lat, max.lat, herring.locations)


```




Summarize number of samples by categories
```{r}

herring.data <- read_csv("herring_diet_basin_data.csv")

atlantis.names <- read_csv("atlantis_groups.csv")


variable.data <- herring.data %>% 
  mutate(index = 1) %>% 
  group_by(variable) %>% 
  summarise(tot_samples = sum(index)) %>% 
  mutate(tot = sum(tot_samples), prop_samples = round((tot_samples/ tot)*100,3))


sample.data <- herring.data %>% 
  mutate(index = 1) %>% 
  group_by(sample_source) %>% 
  summarise(tot_samples = sum(index)) %>% 
  mutate(tot = sum(tot_samples), prop_samples = round((tot_samples/ tot)*100,3))

digestive.categories <- c("bolus", "esophagus", "gastric_lavage", "gizzard", "gut_contents", "regurgitations")

digestive.samples <- sample.data %>% 
  filter(sample_source %in% digestive.categories)

digestive.samples %>% pull(prop_samples) %>% sum()

other.categories <- c("caught_prey","fatty_acids_blubber", "predation_event")

other.samples <- sample.data %>% 
  filter(sample_source %in% other.categories)

other.samples %>% pull(prop_samples) %>% sum()

year.data <- herring.data %>% 
  mutate(index_rows = 1) %>% 
  group_by(atlantis_pred_group, year) %>%
  summarize(tot_data = sum(index_rows)) %>% 
  mutate(yr.fact = as.factor(year)) %>% 
  left_join(atlantis.names, by = c("atlantis_pred_group")) %>% 
  mutate(group_name = if_else(group_name == "Seabirds (mainly non-piscivorous)","Seabirds (non-piscivorous)",
                              if_else(group_name == "Hake /Large gadoids", "Hake & Large gadoids", group_name)))


basin.data <- herring.data %>% 
  mutate(index_rows = 1) %>% 
  group_by(atlantis_pred_group, predator_stage, basin) %>%
  summarize(tot_data = sum(index_rows))

ps.data <- herring.data %>% 
  mutate(index_rows = 1) %>% 
  group_by(atlantis_pred_group, predator_stage) %>%
  summarize(tot_data = sum(index_rows))

herring.data %>% 
  distinct(atlantis_pred_group, predator_taxa) %>% 
  arrange(atlantis_pred_group, predator_taxa)

red.pal1 <- redmonder.pal(8,"qMSOBuWarm")[c(2,1)]

sample.plot <- basin.data %>% 
      ggplot(aes(x=atlantis_pred_group, y=tot_data, fill = predator_stage)) + 
      geom_col()+
      scale_fill_manual(values=red.pal1, name = "Predator stage")+
      theme_classic()+
      theme(legend.position = "bottom")+
      ylab("Data samples") + 
      xlab("Predator group")
 
red.pal2 <-  c(redmonder.pal(8,"qMSOMed"),redmonder.pal(8,"qMSOPap")[8])

basin.plot <- basin.data %>% 
      ggplot(aes(x=atlantis_pred_group, y=tot_data, fill = basin)) + 
      geom_col()+
      scale_fill_manual(values=red.pal2, name = "Basin")+
      theme_classic()+
      theme(legend.position = "bottom")+
      ylab("Data samples") + 
      xlab("Predator group") 

grid.sample <- grid.arrange(sample.plot, basin.plot, ncol=1)

ggsave("herring_samples_plot.png", grid.sample, device="png",width=10,height=10, dpi=350)

   
year.plot <- year.data %>% 
      ggplot(aes(x=year, y=tot_data, group = group_name)) + 
  geom_bar(color = "steelblue4", stat="identity")+
   #   geom_point(size=1.5, color = "steelblue4")+
 # geom_line(color = "steelblue2")+
  facet_wrap(~ group_name)+
     # scale_fill_manual(values=red.pal2, name = "Basin")+
      theme_classic()+
      ylab("Data samples") + 
      xlab("Year") +
  scale_x_continuous(breaks=seq(1970, 2020, 15))
   
ggsave("herring_year_plot.png", year.plot, device="png",width=9,height = 7,dpi=350)
    


  
```

Add covariates to model and plot distribution
```{r}

#used http://jacolienvanrij.com/Tutorials/GAMM.html
#https://fromthebottomoftheheap.net/2014/05/09/modelling-seasonal-data-with-gam/

herring.basin.data <- read_csv("herring_diet_basin_data.csv")

herring.basin.data.scaled <- herring.basin.data %>% 
  mutate(proportion_sc=(0.999*(0.00001+proportion)))
#beta distribution does not use 0 or 1 values
#Ainsworth et al used this approach to scale the diet proportion values
#dat2 <- cbind(x[,1:2],(0.999*(0.00001 + x[,3:(nprey+2)]/rowSums(x[,3:(nprey+2)]))))

herring.spawning <- read_csv("herring_spawning_numbers.csv")

ps.indicators <- read_xlsx("Herring_indicators.xlsx", sheet=3) %>% 
  dplyr::select(Year, SOG.Herring, chinook.ab, chum.ab, coho.ab, pink.ab, sockeye.ab, steelhead.ab, Seals.Nelson, Orcas, Transient_orcas, RR.SSTOut, Chla.JDF, FraserFlow.Out,PNI,NPGO, PDO, UPW, MEI) %>% 
  dplyr::rename(year=Year) %>% 
  left_join(herring.spawning, by="year") %>% 
  dplyr::select(-all_herring) %>% 
   mutate(across(where(is.character),as.numeric)) 

 

ps.indicator.plot <- ps.indicators %>% 
  pivot_longer(-year,names_to="variable") %>% 
  mutate(variable_factor=as.factor(variable))

ps.indicator.plot$variable_factor <- factor(ps.indicator.plot$variable_factor, c("cherry_point", "other_herring", "SOG.Herring",                                               "chinook.ab","chum.ab","coho.ab","pink.ab","steelhead.ab","sockeye.ab",
      "Seals.Nelson","Orcas","Transient_orcas",                                                                                      "FraserFlow.Out","PNI","RR.SSTOut","Chla.JDF","NPGO","PDO","MEI", "UPW"))

col.pal1 <- c(redmonder.pal(8,"qMSOPu"))

rev.pal <- c(rep(col.pal1[8],3),rep(col.pal1[6],9),rep(col.pal1[4],4),rep(col.pal1[3],4))


indicator.plot <- ggplot(data=ps.indicator.plot) +
  geom_line(aes(x=year, y = value, color = variable_factor)) +
  scale_color_manual(values=rev.pal) +
  facet_wrap(~ variable_factor, scales = "free_y", strip.position = "left", 
                labeller = as_labeller(c(cherry_point = "Cherry point herring (mt)",other_herring = "Other herring (mt)",
                                         SOG.Herring="(SOG herring (mt)",chinook.ab = "Chinook (mt)",chum.ab = "Chum (mt)",coho.ab = "Coho (mt)",pink.ab="Pink (mt)",steelhead.ab="Steelhead (mt)",sockeye.ab="Sockeye (mt)",Seals.Nelson="Harbour seals (No.)",Orcas = "Southern Resident orcas (No.)", Transient_orcas = "Transient orcas (No.)",
FraserFlow.Out = "Fraser river flow (m3s-1)",PNI="PNI",RR.SSTOut = "SST (°C)",`Chla.JDF`="Chl a",NPGO="NPGO",PDO="PDO",MEI="MEI", UPW="UPW"))) +
    theme_minimal() +
    theme(legend.position = "none") 
 
  
ggsave("indicator_plot.png", indicator.plot, device="png",width=14,height=8, dpi=250)
    

#https://www.statsandr.com/blog/correlogram-in-r-how-to-highlight-the-most-correlated-variables-in-a-dataset/
#https://datascienceplus.com/find-insights-with-ranked-cross-correlations/

#Ranked cross-correlations

ranked.cross <- ps.indicators %>% 
  setNames(c("year","SOG herring","Chinook", "Chum", "Coho", "Pink", "Sockeye", "Steelhead", "Harbour seals", "Southern Resident orcas", "Transient orcas", "SST", "Chl a", "Fraser river flow", "PNI", "NPGO", "PDO", "UPW", "MEI", "Cherry point herring","Other herring")) %>% 
  dplyr::select(-year) %>% 
  corr_cross(., # name of dataset
  max_pvalue = 0.05, # display only significant correlations (at 5% level)
  top = 20 # display top 10 couples of variables (by correlation coefficient)
)

ggsave("rank_correlations_plot.png", ranked.cross, device="png",width=10,height=9, dpi=300)



herring.data.spawning <- herring.basin.data.scaled %>% 
  left_join(ps.indicators, by="year")

ps.basins <- herring.data.spawning %>% 
  distinct(basin) %>% 
  arrange(basin) %>% 
  mutate(basin_id=1:nrow(.))

pred.stages <- herring.data.spawning %>% 
  distinct(predator_stage) %>% 
  arrange(predator_stage) %>% 
  mutate(predator_stage_id=1:nrow(.))

pred.groups <- herring.data.spawning %>% 
  distinct(atlantis_pred_group) %>% 
  arrange(atlantis_pred_group) %>% 
  mutate(atlantis_pred_group_id=1:nrow(.))

herring.basin.data.mod <- herring.data.spawning %>% 
  left_join(ps.basins, by="basin") %>% 
  left_join(pred.stages, by="predator_stage") %>% 
  left_join(pred.groups, by="atlantis_pred_group") 

write_csv(herring.basin.data.mod, "herring_diet_basin_data_mod.csv")

#calculate multicollinearity
#http://www.sthda.com/english/articles/39-regression-model-diagnostics/160-multicollinearity-essentials-and-vif-in-r/

ps.indicator.lm <-  ps.indicators %>% 
   setNames(c("year","SOG herring","Chinook", "Chum", "Coho", "Pink", "Sockeye", "Steelhead", "Harbour seals", "Southern Resident orcas", "Transient orcas", "SST", "Chl a", "Fraser river flow", "PNI", "NPGO", "PDO", "UPW", "MEI", "Cherry point herring","Other herring")) %>% 
   filter(year!=2016) %>% 
  dplyr::select(-`Transient orcas`)

model1 <- lm(year ~ ., data = ps.indicator.lm, na.action = na.exclude)

vif.res <- car::vif(model1) %>% 
  as.data.frame() %>% 
  as_tibble(rownames = "variable") %>% 
  setNames(c("variable","VIF")) %>% 
  mutate(variable_factor = as.factor(variable))


vif.res$variable_factor <- factor(vif.res$variable_factor, c("`Cherry point herring`","`Other herring`", "`SOG herring`","Chinook", "Chum", "Coho", "Pink", "Steelhead", "Sockeye", "`Harbour seals`", "`Southern Resident orcas`", "`Fraser river flow`", "PNI", "SST", "`Chl a`", "NPGO", "PDO", "MEI", "UPW")) 

col.pal2 <- c(redmonder.pal(9,"sPBIGy1")[3:5],redmonder.pal(9,"sPBIBu")[2:9],redmonder.pal(9,"sPBIPu")[4:7],redmonder.pal(9,"sPBIRdPu")[2:5])

vif.plot <- ggplot(aes(x=variable_factor,y=VIF, fill=variable_factor), data=vif.res) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=col.pal2, name="Variables", labels = c("Cherry point herring (mt)", "Other herring (mt)", "SOG herring (mt)", "Chinook (mt)", "Chum (mt)", "Coho (mt)","Pink (mt)","Steelhead (mt)","Sockeye (mt)","Harbour seals (No.)", "Southern Resident Orcas (No.)",
"Fraser river flow (m3s-1)","PNI","SST (°C)","Chl a","NPGO","PDO","MEI","UPW"))+
   theme(axis.text.x = element_blank())+
  labs(x="Variable")+
  geom_hline(yintercept=3)

ggsave("vif_multicolinearity_plot.png", vif.plot, device="png",width=10,height=8, dpi=200)

write_csv(vif.res,"VIF_values.csv")

# examine data
# see https://cran.r-project.org/web/packages/fitdistrplus/vignettes/FAQ.html#can-i-personalize-the-default-plot-given-for-an-object-of-class-fitdist-or-fitdistcens

fit.norm <- fitdist(herring.basin.data.mod$proportion_sc, distr = "beta", method = "mme")
plot(fit.norm)


dens.plot <- denscomp(fit.norm, addlegend = FALSE, main = "", xlab = "Herring prey contribution", ylim=c(0,10), fitcol = "red", plotstyle = "ggplot") +
  ggplot2::theme_classic()
qq.plot <- qqcomp(fit.norm, addlegend = FALSE, main = "", fitpch = 16, fitcol = "grey", line01lty = 2, plotstyle = "ggplot") +
  ggplot2::theme_classic()
cdf.plot <- cdfcomp(fit.norm, addlegend = FALSE, main = "", xlab = "Herring prey contribution", fitcol = "red", lines01 = TRUE, plotstyle = "ggplot") +
  ggplot2::theme_classic()
pp.plot <- ppcomp(fit.norm, addlegend = FALSE, main = "", fitpch = 16, fitcol = "grey", line01lty = 2, plotstyle = "ggplot") +
  ggplot2::theme_classic()

myplot1 <- arrangeGrob(dens.plot, top = textGrob("A", x = unit(0, "npc")
         , y   = unit(1, "npc"), just=c("left","top"),
         gp=gpar(col="black", fontsize=14, fontfamily="Arial")))

myplot2 <- arrangeGrob(qq.plot, top = textGrob("B", x = unit(0, "npc")
         , y = unit(1, "npc"), just=c("left","top"),
         gp=gpar(col="black", fontsize=14, fontfamily="Arial")))

myplot3 <- arrangeGrob(cdf.plot, top = textGrob("C", x = unit(0, "npc")
        , y  = unit(1, "npc"), just=c("left","top"),
        gp=gpar(col="black", fontsize=14, fontfamily="Arial")))

myplot4 <- arrangeGrob(pp.plot, top = textGrob("D", x = unit(0, "npc")
        , y = unit(1, "npc"), just=c("left","top"),
        gp=gpar(col="black",    fontsize=14, fontfamily="Arial")))

grid.sample <- grid.arrange(myplot1, myplot2, myplot3, myplot4, ncol=2, nrow=2)

ggsave("distribution.fit.png", grid.sample, device="png",width=10,height=10, dpi=300)



```



PCA for environmental variables

```{r}


herring.basin.data.mod <- read_csv("herring_diet_basin_data_mod.csv") %>% 
 mutate(atlantis_pred_group = as.factor(atlantis_pred_group),
        basin=as.factor(basin)) %>% 
  mutate(rec_index=1:nrow(.)) 

#pull out SOG.Herring, cherry_point, other_herring
#`SOG herring` = SOG.Herring, `Cherry Point herring` = cherry_point, `Other herring` = other_herring,
herring.environ <- herring.basin.data.mod %>% 
  dplyr::select(chinook.ab, chum.ab, coho.ab, pink.ab, sockeye.ab, steelhead.ab, Seals.Nelson, Orcas, RR.SSTOut, Chla.JDF, FraserFlow.Out,PNI,NPGO, PDO, UPW, MEI) %>% 
  dplyr::rename(Chinook = chinook.ab, Chum = chum.ab, Coho = coho.ab, Pink = pink.ab, Sockeye = sockeye.ab, Steelhead = steelhead.ab, `Southern Resident orcas`= Orcas, `Harbour seals` = Seals.Nelson, SST = RR.SSTOut, `Chl a` = Chla.JDF, `Fraser river flow`=FraserFlow.Out)


#PCA resources
#http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/
#https://aaronschlegel.me/principal-component-analysis-r-example.html
#http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/
#https://github.com/rogih/GAM-PCA-VAR/blob/master/RD_GAM_PCA_VAR.R
#https://uc-r.github.io/pca#selecting

#standarizing data
stand.environ.index <- vegan::decostand(herring.environ, "standardize") %>% 
  mutate(rec_index=1:nrow(.)) %>% 
  na.omit

stand.environ <- vegan::decostand(herring.environ, "standardize") %>% 
  na.omit 

  # Do the PCA 
#To perform principal component analysis using the correlation matrix using the prcomp() function, set the scale argument to TRUE.
#This approach is recommended for data that is measured in different units or has wide variances.

res.pca <- prcomp(stand.environ,center=T,scale=T)

res.cov <- cov(stand.environ)
res.eigen <- eigen(res.cov)

PVE <- res.eigen$values / sum(res.eigen$values)

#attach pc axes later selected to data
pca.axes <- res.pca$x %>% 
  as_tibble %>% 
  bind_cols(stand.environ.index) %>% 
  dplyr::select(PC1,PC2,PC3,PC4,PC5,PC6, rec_index) %>% 
  left_join(herring.basin.data.mod, by = "rec_index")
  
write_csv(pca.axes, "herring_diet_basin_data_pca.csv")

#choose how many dimensions you need
stand.environ.paran <- paran(stand.environ, graph = TRUE)

parallel.analysis <- tibble(AdjustedEv=stand.environ.paran$AdjEv, RandomEv= stand.environ.paran$RndEv, UnadjustedEv=stand.environ.paran$Ev, PVE=PVE) %>% 
  mutate(components=1:nrow(.)) %>% 
  pivot_longer(cols=AdjustedEv:PVE, names_to = "parameter", values_to="eigenvalue") 

col.pal1 <- c(redmonder.pal(8,"qMSOSlp"))
col.pal1 <- col.pal1[c(1,3,4)]

paralell.plot <- parallel.analysis %>% 
  filter(parameter!="PVE") %>% 
  filter(components < 11) %>% 
  mutate(components = as.factor(components)) %>% 
  ggplot(aes(x=components,y=eigenvalue, colour=parameter, group=parameter))+
  geom_line()+
  scale_colour_manual(values=col.pal1, name="Parameter", labels=c("Adjusted Ev","Random Ev","Unadjusted Ev")) +
 # geom_hline(yintercept=1) +
  theme_minimal() +
  xlab("Dimensions") +
  ylab("Eigenvalue") +
  ggtitle("Parallel analysis")


#scree plot
#the plot of eigenvalues ordered from largest to the smallest. The number of component is determined at the point, beyond which the remaining eigenvalues are all relatively small and of comparable size (Jollife 2002, Peres-Neto, Jackson, and Somers (2005)).
scree.plot <- fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50)) +
  theme_minimal()

pca.test.plot <- grid.arrange(paralell.plot,scree.plot, nrow=2, ncol=1)

ggsave("dimensions_plot_pca.png", pca.test.plot, device="png",width=8,height = 6,dpi=350)



#graph of variables
var.pca.plot <- fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

ggsave("var_plot_pca.png", var.pca.plot, device="png",width=10,height = 8,dpi=350)


# #graph of individual data points
# fviz_pca_ind(res.pca,
#              col.ind = "cos2", # Color by the quality of representation
#              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#              repel = TRUE     # Avoid text overlapping
#              )

pca.plot <- autoplot(res.pca, data = stand.environ)
pca.plot

# Eigenvalues
#http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/
# examine the eigenvalues to determine the number of principal components to be considered. 

#An eigenvalue > 1 indicates that PCs account for more variance than accounted by one of the original variables in standardized data. This is commonly used as a cutoff point for which PCs are retained. This holds true only when the data are standardized.

#You can also limit the number of component to that number that accounts for a certain fraction of the total variance. For example, if you are satisfied with 70% of the total variance explained then use the number of components to achieve that.

eig.val <- get_eigenvalue(res.pca)
eig.val
  
# Results for Variables
res.var <- get_pca_var(res.pca)
res.var$coord          # Coordinates
res.var$contrib        # Contributions to the PCs
res.var$cos2           # Quality of representation 


#The quality of representation of the variables on factor map is called cos2 (square cosine, squared coordinates)

png(height=800, width=800, file="corr_plot_pca.png", type = "cairo")
corrplot(res.var$cos2, 
         method = "circle",
         is.corr = F,
         order = "original",
         tl.col="black")
dev.off()



# Contributions of variables to PC1
pca.comp1 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)+
  ggtitle("Dim 1")
pca.comp2 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)+
  ggtitle("Dim 2")
pca.comp3 <- fviz_contrib(res.pca, choice = "var", axes = 3, top = 10)+
  ggtitle("Dim 3")
pca.comp4 <- fviz_contrib(res.pca, choice = "var", axes = 4, top = 10)+
  ggtitle("Dim 4")
pca.comp5 <- fviz_contrib(res.pca, choice = "var", axes = 5, top = 10)+
  ggtitle("Dim 5")
pca.comp6 <- fviz_contrib(res.pca, choice = "var", axes = 6, top = 10)+
  ggtitle("Dim 6")

var.contribution <- grid.arrange(pca.comp1,pca.comp2,pca.comp3,pca.comp4,pca.comp5,pca.comp6, nrow=3, ncol = 2)

# Total cos2 of variables on Dim.1 to Dim.4
fviz_cos2(res.pca, choice = "var", axes = 1)

# A high cos2 indicates a good representation of the variable on the principal component. In this case the variable is positioned close to the circumference of the correlation circle.
# 
# A low cos2 indicates that the variable is not perfectly represented by the PCs. In this case the variable is close to the center of the circle.



```




GAM model over PCA 

```{r}

#GAM resources
#https://m-clark.github.io/generalized-additive-models/application.html#multiple-predictors

herring.basin.data.pc <- read_csv("herring_diet_basin_data_pca.csv") %>% 
  mutate(basin_id = as.factor(basin_id), atlantis_pred_group = as.factor(atlantis_pred_group),
        basin=as.factor(basin)) 


#gamma 1.4 to reduce overfitting

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

#full model with no interaction terms
base="tp"

herring.m1 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, bs=base, k = 12) +
                    s(atlantis_pred_group, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
            #      s(predator_stage_id, bs = base, k = 1) + # not used because k was too low
                    s(PC1, bs = base, k=8) +
                   s(PC2, bs=base, k=8) +
                   PC3 + #edf ~1 when smooth term, used as parametric
                    s(PC4, bs=base, k=8) +
                    s(PC5, bs = base, k=8) +
                   s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m1,"herring_gam_pc_m1.rds")

gam.check(herring.m1)

try(concurvity(herring.m1, full = TRUE))

model.stats.m1 <- tidy(herring.m1)

model.stats.m1

#gamma 1.4 to reduce overfitting

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

base="tp"

herring.m2 <- gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(predator_size, by=year, bs=base, k = 12) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(basin, by=year, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                    SOG.Herring +
                    s(cherry_point, bs=base, k=10) +
                   s(other_herring, bs=base, k=10) +
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(PC1, bs = base, k=8) +
                    s(PC2, bs=base, k=8) +
                   PC3 +
                    s(PC4, bs=base, k=8) +
                    s(PC5, bs = base, k=8) +
                    s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m2,"herring_gam_pc_m2.rds")

gam.check(herring.m2)

try(concurvity(herring.m2, full = TRUE))

model.stats.m2 <- tidy(herring.m2)

model.stats.m2


#by basin

base="tp"

herring.m3 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    predator_size * basin +
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    s(year, by=basin, bs=base, k=10) +
                    s(basin, by=year, bs='re', k=10) +
                    s(SOG.Herring, bs=base, k= 10) +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(PC1, bs = base, k=8) +
                    s(PC2, bs= base, k=8) +
                    s(PC3, bs= base, k=8) +
                    s(PC4, bs= base, k=8) +
                    s(PC5, bs = base, k=8) +
                    s(PC6, bs= base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m3,"herring_gam_pc_m3.rds")

gam.check(herring.m3)

try(concurvity(herring.m3, full = TRUE))

model.stats.m3 <- tidy(herring.m3)

model.stats.m3



#combines year and basin
base="tp"

herring.m4 <- gam(proportion_sc ~ 
                   s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    predator_size * basin +
                    s(year, by=basin, bs=base, k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    basin +
                    s(year, bs=base, k=10) +
                    s(SOG.Herring, bs=base, k= 10) +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
                    s(PC1, bs = base, k=8) +
                    s(PC2, bs=base, k=8) +
                    s(PC3, bs=base, k=8) +
                    s(PC4, bs=base, k=8) +
                    s(PC5, bs= base, k=8) +
                    s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)


saveRDS(herring.m4,"herring_gam_pc_m4.rds")

gam.check(herring.m4)

try(concurvity(herring.m4, full = TRUE))

model.stats.m4 <- tidy(herring.m4)

model.stats.m4


#ts
#full model with no interaction terms
base="ts"

herring.m5 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, bs=base, k = 12) +
                    s(atlantis_pred_group, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                    s(SOG.Herring, bs=base, k= 10) +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(PC1, bs = base, k=8) +
                    s(PC2, bs=base, k=8) +
                    s(PC3, bs=base, k=8) +
                    s(PC4, bs=base, k=8) +
                    s(PC5, bs = base, k=8) +
                    s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m5,"herring_gam_pc_m5.rds")

gam.check(herring.m5)

try(concurvity(herring.m5, full = TRUE))

model.stats.m5 <- tidy(herring.m5)

model.stats.m5


#gamma 1.4 to reduce overfitting

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

base="ts"

herring.m6 <- gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(predator_size, by=year, bs=base, k = 12) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    basin * year +
                    s(year, bs=base, k=10) +
                    SOG.Herring +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
                   # s(predator_stage_id, bs = 'ts', k = 1) +
                    s(PC1, bs = base, k=8) +
                    s(PC2, bs=base, k=8) +
                    PC3 +
                    s(PC4, bs=base, k=8) +
                    s(PC5, bs = base, k=8) +
                    s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m6,"herring_gam_pc_m6.rds")

gam.check(herring.m6)

try(concurvity(herring.m6, full = TRUE))

model.stats.m6 <- tidy(herring.m6)

model.stats.m6



base="ts"

herring.m7 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, by=basin, bs=base, k = 12) +
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    year * basin +
                    s(basin, bs='re', k=10) +
                    s(SOG.Herring, bs=base, k=10) +
                    s(cherry_point, bs=base, k=10) +
                   s(other_herring, bs=base, k=10) +
                  # s(predator_stage_id, bs = 'ts', k = 1) +
                   s(PC1, bs = base, k=8) +
                    s(PC2, bs=base, k=8) +
                   s(PC3, bs=base, k=8) +
                   s(PC4, bs=base, k=8) +
                   s(PC5, bs = base, k=8) +
                    s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m7,"herring_gam_pc_m7.rds")

gam.check(herring.m7)

try(concurvity(herring.m7, full = TRUE))

model.stats.m7 <- tidy(herring.m7)

model.stats.m7
#combines year and basin
base="ts"

herring.m8 <- gam(proportion_sc ~ 
                   s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    s(predator_size, by=basin, bs=base, k = 12) +
                    year * basin +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(basin, bs = 're', k=10) +
                    s(year, bs=base, k=10) +
                    s(SOG.Herring, bs=base, k=10) +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
                   s(PC1, bs = base, k=8) +
                   s(PC2, bs=base, k=8) +
                  s(PC3, bs=base, k=8) +
                   s(PC4, bs=base, k=8) +
                   s(PC5, bs = base, k=8) +
                    s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)


saveRDS(herring.m8,"herring_gam_pc_m8.rds")

gam.check(herring.m8)

try(concurvity(herring.m8, full = TRUE))

model.stats.m8 <- tidy(herring.m8)

model.stats.m8

#added later, so it first does tp and then ts
#by atlantis_pred group
base="tp"

herring.m9 <- gam(proportion_sc ~ 
                   s(longitude, latitude, bs=base, k = 100) + 
                    s(year, bs=base, k=10) +
                    s(predator_size, bs=base, k = 12) +
                    s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                    s(atlantis_pred_group, bs='re', k = 10) +
                    s(SOG.Herring, bs=base, k=10) +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
                    s(PC1, bs = base, k=8) +
                    s(PC2, bs=base, k=8) +
                    s(PC3, bs=base, k=8) +
                    s(PC4, bs=base, k=8) +
                    PC5 +
                    s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)


saveRDS(herring.m9,"herring_gam_pc_m9.rds")

gam.check(herring.m9)

try(concurvity(herring.m9, full = TRUE))

model.stats.m9 <- tidy(herring.m9)

model.stats.m9



base="ts"

herring.m10 <- gam(proportion_sc ~ 
                   s(longitude, latitude, bs=base, k = 100) + 
                    s(year, bs=base, k=10) +
                    s(predator_size, bs=base, k = 12) +
                    s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                    s(atlantis_pred_group, bs='re', k = 10) +
                   s(SOG.Herring, bs=base, k=10) +
                    s(cherry_point, bs=base, k=10) +
                    s(other_herring, bs=base, k=10) +
                    s(PC1, bs = base, k=8) +
                    s(PC2, bs=base, k=8) +
                    s(PC3, bs=base, k=8) +
                    s(PC4, bs=base, k=8)  +
                   s(PC5, bs = base, k=8) +
                   s(PC6, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m10,"herring_gam_pc_m10.rds")

gam.check(herring.m10)

try(concurvity(herring.m10, full = TRUE))

model.stats.m10 <- tidy(herring.m10)

model.stats.m10



base="ts"

herring.m11 <- gam(proportion_sc ~ 
                        s(longitude, latitude, by=year, bs=base, k = 100) + 
                        s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                        s(predator_size, by=year, bs=base, k = 10) +
                        s(year, by=basin, bs=base, k=10) +
                        s(atlantis_pred_group, by=year, bs = 're', k=10) +
                        s(year, by=atlantis_pred_group, bs = base, k=10) + 
                        basin +
                        year +
                        s(year, bs=base, k=10) +
                        s(SOG.Herring, bs=base, k= 10) +
                        s(cherry_point, bs=base, k=10) +
                        s(other_herring, bs=base, k=10) +
                        s(PC1, bs = base, k=8) +
                        s(PC2, bs=base, k=8) +
                        s(PC3, bs=base, k=8) +
                        s(PC4, bs=base, k=8) +
                        s(PC5, bs= base, k=8) +
                        s(PC6, bs=base, k=8),
                    family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)

saveRDS(herring.m11,"herring_gam_pc_m11.rds")

gam.check(herring.m11)

try(concurvity(herring.m11, full = TRUE))

model.stats.m11 <- tidy(herring.m11)

model.stats.m11


base="tp"

herring.m12 <- gam(proportion_sc ~ 
                        s(longitude, latitude, by=year, bs=base, k = 100) + 
                        s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                        s(predator_size, by=year, bs=base, k = 10) +
                        s(year, by=basin, bs=base, k=10) +
                        s(atlantis_pred_group, by=year, bs = 're', k=10) +
                        s(year, by=atlantis_pred_group, bs = base, k=10) + 
                        basin +
                        year +
                        s(year, bs=base, k=10) +
                        s(SOG.Herring, bs=base, k= 10) +
                        s(cherry_point, bs=base, k=10) +
                        s(other_herring, bs=base, k=10) +
                        s(PC1, bs = base, k=8) +
                        s(PC2, bs=base, k=8) +
                        s(PC3, bs=base, k=8) +
                        s(PC4, bs=base, k=8) +
                        s(PC5, bs= base, k=8) +
                        s(PC6, bs=base, k=8),
                    family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.pc)


saveRDS(herring.m12,"herring_gam_pc_m12.rds")

gam.check(herring.m12)

try(concurvity(herring.m12, full = TRUE))

model.stats.m12 <- tidy(herring.m12)

model.stats.m12



herring_gam_pc_m1 <- readRDS("~/herringdiet/herring_gam_pc_m1.rds")
herring_gam_pc_m2 <- readRDS("~/herringdiet/herring_gam_pc_m2.rds")
herring_gam_pc_m3 <- readRDS("~/herringdiet/herring_gam_pc_m3.rds")
herring_gam_pc_m4 <- readRDS("~/herringdiet/herring_gam_pc_m4.rds")
herring_gam_pc_m5 <- readRDS("~/herringdiet/herring_gam_pc_m5.rds")
herring_gam_pc_m6 <- readRDS("~/herringdiet/herring_gam_pc_m6.rds")
herring_gam_pc_m7 <- readRDS("~/herringdiet/herring_gam_pc_m7.rds")
herring_gam_pc_m8 <- readRDS("~/herringdiet/herring_gam_pc_m8.rds")
herring_gam_pc_m9 <- readRDS("~/herringdiet/herring_gam_pc_m9.rds")
herring_gam_pc_m10 <- readRDS("~/herringdiet/herring_gam_pc_m10.rds")
herring_gam_pc_m11 <- readRDS("~/herringdiet/herring_gam_pc_m11.rds")
herring_gam_pc_m12 <- readRDS("~/herringdiet/herring_gam_pc_m12.rds")

list.fits <- list(herring_gam_pc_m1,herring_gam_pc_m2,herring_gam_pc_m3,herring_gam_pc_m4,herring_gam_pc_m5,
                  herring_gam_pc_m6,herring_gam_pc_m7,herring_gam_pc_m8, herring_gam_pc_m9, herring_gam_pc_m10, herring_gam_pc_m11, herring_gam_pc_m12)

no.fits <- 1:length(list.fits)

gam.res.table <- lapply(no.fits, gam_table, list.fits) %>% 
  bind_rows %>% 
  arrange(desc(AIC))

write_csv(gam.res.table, "GAM_table_res_pca.csv")


#Final model

#model 

herring.pc.final <- herring_gam_pc_m12

saveRDS(herring.pc.final,"herring_gam_pc_final.rds")




#https://cran.r-project.org/web/packages/mgcViz/vignettes/mgcviz.html
herring.viz <- getViz(herring.pc.final)

print(plot(herring.viz, allTerms = T), pages = 1) # Calls print.plotGam()

#herring.pc.draw <- draw(herring.pc.final)+
#  theme_classic()

herring.pc.appraise <- appraise(herring.pc.final)

#use to plot 2D
#plot(sm(herring.viz, 2)) + l_fitRaster() + l_fitContour() + l_points()

ggsave("pc_gam_appraise.png", herring.pc.appraise, device="png",width=10,height=8, dpi=350)

#ggsave("pc_gam_draw.png", herring.pc.draw, device="png",width=10,height=8, dpi=350)


#use sm to extract smooth terms from model

herring.viz.term.1 <- plot(sm(herring.viz, 7)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "A", x="Gaussian quantiles", y = "Effects", title="s(basin):predator group", subtitle = "Dogfish")

herring.viz.term.2 <- plot(sm(herring.viz, 11)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "B", x="Gaussian quantiles", y = "Effects", title="s(basin):predator group", subtitle = "Large demersal predators")

herring.viz.term.3 <- plot(sm(herring.viz, 18)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "C", x="Gaussian quantiles", y = "Effects", title="s(basin):predator group", subtitle = "Orcas")


herring.viz.term.4 <- plot(sm(herring.viz, 25)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "D", x="Predator size", y = "s(predator size):year", title="", subtitle = "")

herring.viz.term.5 <- plot(sm(herring.viz, 29)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "E", x="Year", y = "Effects", title="s(year):basin", subtitle = "San Juan Islands")

herring.viz.term.6 <- plot(sm(herring.viz, 40)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "F", x="Year", y = "s(year):predator group", title="", subtitle = "Dogfish")

herring.viz.term.7 <- plot(sm(herring.viz, 44)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "G", x="Year", y = "s(year):predator group", title="", subtitle = "Large demersal predators")

herring.viz.term.8 <- plot(sm(herring.viz, 51)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "H", x="Year", y = "s(year):predator group", title="", subtitle = "Orcas")


# herring.viz.term.8 <- plot(pterm(herring.viz, 1)) +
#   l_ciBar(colour = "darkblue") + 
#   l_fitPoints(colour = "darkred") + 
#            l_rug(alpha = 0.3) +
#   #  l_fitLine(colour = "black") + 
#   #l_rug(mapping = aes(x=x, y=y), alpha = 0.8) +
#   #  l_ciLine(mul = 5, colour = "blue", linetype = 2) + 
#   #  l_points(shape = 19, size = 1, alpha = 0.1) + 
#   labs(tag = "H") +
#   labs(x="Predator group", y = "f(Predator group)")+
#   ggtitle("")+
#   scale_x_discrete(guide = guide_axis(n.dodge=3))+
#   theme_classic() +
#   theme(axis.text.x = element_text(size=6))




grid.sample <- gridPrint(herring.viz.term.1, herring.viz.term.2, herring.viz.term.3,
                         herring.viz.term.4, herring.viz.term.5, herring.viz.term.6, 
                         herring.viz.term.7, herring.viz.term.8, 
                         ncol=3, nrow=3)


ggsave("pc_gam_plot.png", grid.sample, device="png",width=10,height=6, dpi=350)


```

All species, with all environmental variables


```{r}

herring.basin.data.mod <- read_csv("herring_diet_basin_data_mod.csv") %>% 
 mutate(basin_id = as.factor(basin_id), atlantis_pred_group = as.factor(atlantis_pred_group),
        basin=as.factor(basin)) 

pred.year <- herring.basin.data.mod %>% 
  distinct(year,atlantis_pred_group) %>% 
  arrange(atlantis_pred_group,year)

pred.year %>% pull(year)

# examine data
# see https://cran.r-project.org/web/packages/fitdistrplus/vignettes/FAQ.html#can-i-personalize-the-default-plot-given-for-an-object-of-class-fitdist-or-fitdistcens

#gamma 1.4 to reduce overfitting

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

#full model with no interaction terms
base="tp"

herring.m1 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, bs=base, k=10) +
                    s(atlantis_pred_group, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) + 
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

saveRDS(herring.m1,"herring_gam_vars_m1.rds")

gam.check(herring.m1)

concurvity(herring.m1, full = TRUE)

model.stats.all.m1 <- tidy(herring.m1) 

#model with interaction terms by year

base="tp"

herring.m2 <- gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(predator_size, by=year, bs=base, k = 10) + 
                    s(basin, by=year, bs='re', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(year, bs=base, k = 10) +
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) + 
                    s(Seals.Nelson, bs=base, k=8) + 
                    #s(Orcas, bs=base, k=8) +
                    Orcas +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs = base, k=8) + +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


saveRDS(herring.m2,"herring_gam_vars_m2.rds")

gam.check(herring.m2)

model.stats.all.m2 <- tidy(herring.m2) 


#model with interaction terms by basin

base="tp"

herring.m3 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(year, by=basin, bs=base, k=10) +
                    s(atlantis_pred_group, by=basin, bs='re', k=10) +
                    s(basin, bs='re', k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) + 
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


#concurvity(herring.m3, full = TRUE)

saveRDS(herring.m3,"herring_gam_vars_m3.rds")

gam.check(herring.m3)

model.stats.all.m3 <- tidy(herring.m3) 


#combines year and basin
base="tp"

herring.m4 <- gam(proportion_sc ~ 
                   s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    s(year, by=basin, bs=base, k=10) +
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                   s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) + 
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

#concurvity(herring.m4, full = TRUE)

saveRDS(herring.m4,"herring_gam_vars_m4.rds")

gam.check(herring.m4)

model.stats.all.m4 <- tidy(herring.m4) 


#by atlantis_pred_group
base="tp"

herring.m5 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(year, bs=base, k=10) +
                    s(predator_size, bs=base, k=10) + 
                    s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) +
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

#concurvity(herring.m5, full = TRUE)

saveRDS(herring.m5,"herring_gam_vars_m5.rds")

gam.check(herring.m5)

model.stats.all.m5 <- tidy(herring.m5) 



#same as models 1-5 but uses ts as a basis

base="ts"

herring.m6 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, bs=base, k = 12) +
                    s(atlantis_pred_group, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                   s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                   # SOG.Herring +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    #coho.ab +
                    s(sockeye.ab, bs=base, k=8) +
                    #sockeye.ab +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) + 
                    #pink.ab +
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    #Chla.JDF +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


saveRDS(herring.m6,"herring_gam_vars_m6.rds")

gam.check(herring.m6)

model.stats.all.m6 <- tidy(herring.m6) 


#concurvity(herring.m5, full = TRUE)

#model with interaction terms by year

base="ts"

herring.m7 <- gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    #predator_size*year + 
                    s(predator_size, by=year, bs=base, k=10) +
                    s(basin, by=year, bs='re', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(year, bs=base, k=10) +
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    #other_herring +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs=base, k=8) +
                    #chum.ab +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                   # coho.ab +
                    s(sockeye.ab, bs=base, k=8) +
                    #sockeye.ab +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) +
                   # pink.ab +
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    #Chla.JDF +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


saveRDS(herring.m7,"herring_gam_vars_m7.rds")

gam.check(herring.m7)

model.stats.all.m7 <- tidy(herring.m7) 

#model with interaction terms by basin

base="ts"

herring.m8 <- gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(year, by=basin, bs=base, k=10) +
                    s(atlantis_pred_group, by=basin, bs='re', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) + 
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

#concurvity(herring.m7, full = TRUE)

saveRDS(herring.m8,"herring_gam_vars_m8.rds")

gam.check(herring.m8)

model.stats.all.m8 <- tidy(herring.m8) 


#combines year and basin
base="ts"

herring.m9 <- gam(proportion_sc ~ 
                   s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    s(year, by=basin, bs=base, k=10) +
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                   s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) + 
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


try(concurvity(herring.m9, full = TRUE))

saveRDS(herring.m9,"herring_gam_vars_m9.rds")

gam.check(herring.m9)

model.stats.all.m9 <- tidy(herring.m9) 



base="ts"

herring.m10 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(year, bs=base, k=10) +
                    s(predator_size, bs=base, k=10) + 
                    s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) +
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


#concurvity(herring.m5, full = TRUE)

saveRDS(herring.m10,"herring_gam_vars_m10.rds")

gam.check(herring.m10)

model.stats.all.m10 <- tidy(herring.m10) 


base="ts"

herring.m11 <- gam(proportion_sc ~ 
                        s(longitude, latitude, by=year, bs=base, k = 100) + 
                        s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                        s(predator_size, by=year, bs=base, k = 10) +
                        s(year, by=basin, bs=base, k=10) +
                        s(atlantis_pred_group, by=year, bs = 're', k=10) +
                        s(year, by=atlantis_pred_group, bs = base, k=10) + 
                        basin +
                        year +
                     s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) +
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)



saveRDS(herring.m11,"herring_gam_vars_m11.rds")

gam.check(herring.m11)
      
model.stats.all.m11 <- tidy(herring.m11) 


base="tp"

herring.m12 <- gam(proportion_sc ~ 
                        s(longitude, latitude, by=year, bs=base, k = 100) + 
                        s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                        s(predator_size, by=year, bs=base, k = 10) +
                        s(year, by=basin, bs=base, k=10) +
                        s(atlantis_pred_group, by=year, bs = 're', k=10) +
                        s(year, by=atlantis_pred_group, bs = base, k=10) + 
                        basin +
                        year +
                        s(year, bs=base, k=10) +
                       s(other_herring, bs=base, k=8) +
                    s(cherry_point, bs=base, k=8) + 
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(steelhead.ab, bs=base, k=8) + 
                    s(pink.ab, bs=base, k=8) +
                    s(Seals.Nelson, bs=base, k=8) + 
                    s(Orcas, bs=base, k=8) +
                    s(PDO, bs=base, k=8) + 
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

saveRDS(herring.m12,"herring_gam_vars_m12.rds")

gam.check(herring.m12)
      
model.stats.all.m12 <- tidy(herring.m12) 


herring_gam_vars_m1 <- readRDS("~/herringdiet/herring_gam_vars_m1.rds")
herring_gam_vars_m2 <- readRDS("~/herringdiet/herring_gam_vars_m2.rds")
herring_gam_vars_m3 <- readRDS("~/herringdiet/herring_gam_vars_m3.rds")
herring_gam_vars_m4 <- readRDS("~/herringdiet/herring_gam_vars_m4.rds")
herring_gam_vars_m5 <- readRDS("~/herringdiet/herring_gam_vars_m5.rds")
herring_gam_vars_m6 <- readRDS("~/herringdiet/herring_gam_vars_m6.rds")
herring_gam_vars_m7 <- readRDS("~/herringdiet/herring_gam_vars_m7.rds")
herring_gam_vars_m8 <- readRDS("~/herringdiet/herring_gam_vars_m8.rds")
herring_gam_vars_m9 <- readRDS("~/herringdiet/herring_gam_vars_m9.rds")
herring_gam_vars_m10 <- readRDS("~/herringdiet/herring_gam_vars_m10.rds")
herring_gam_vars_m11 <- readRDS("~/herringdiet/herring_gam_vars_m11.rds")
herring_gam_vars_m12 <- readRDS("~/herringdiet/herring_gam_vars_m12.rds")


list.fits <- list(herring_gam_vars_m1,herring_gam_vars_m2,herring_gam_vars_m3,herring_gam_vars_m4,herring_gam_vars_m5,herring_gam_vars_m6,herring_gam_vars_m7,herring_gam_vars_m8, herring_gam_vars_m9, 
                  herring_gam_vars_m10, herring_gam_vars_m11, herring_gam_vars_m12)

model.names <- paste0("model_",1:length(list.fits))
no.fits <- 1:length(list.fits)

#aictab(cand.set = list.fits, modnames = model.names)

gam.res.table <- lapply(no.fits, gam_table, list.fits) %>% 
  bind_rows

write_csv(gam.res.table, "GAM_table_res_vars.csv")

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

#model 1

base="ts"

herring.m.final <- herring_gam_vars_m10
  
saveRDS(herring.m.final,"herring_gam_vars_final.rds")


#https://cran.r-project.org/web/packages/mgcViz/vignettes/mgcviz.html
herring.viz <- getViz(herring.m.final)

print(plot(herring.viz, allTerms = T), pages = 1) # Calls print.plotGam()

#herring.m.draw <- draw(herring.m.final)+
#  theme_classic()

herring.m.appraise <- appraise(herring.m.final)

#use to plot 2D
#plot(sm(herring.viz, 2)) + l_fitRaster() + l_fitContour() + l_points()

ggsave("vars_gam_appraise.png", herring.m.appraise, device="png",width=10,height=8, dpi=350)


#use sm to extract smooth terms from model

herring.viz.term.1 <- plot(sm(herring.viz, 4)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "A", x="Gaussian quantiles", y = "Effects", title = "s(basin): predator group", subtitle = "Chinook")

herring.viz.term.2 <- plot(sm(herring.viz, 5)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
 labs(tag = "A", x="Gaussian quantiles", y = "Effects", title = "s(basin): predator group", subtitle = "Hatchery Chinook")

herring.viz.term.3 <- plot(sm(herring.viz, 9)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "A", x="Gaussian quantiles", y = "Effects", title = "s(basin): predator group", subtitle = "Dogfish")

herring.viz.term.4 <- plot(sm(herring.viz, 13)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
   labs(tag = "A", x="Gaussian quantiles", y = "Effects", title = "s(basin): predator group", subtitle = "Large demersal fish")

herring.viz.term.5 <- plot(sm(herring.viz, 32)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "C", x="Chinook abundance", y = "s(Chinook abundance)")
  
herring.viz.term.6 <- plot(sm(herring.viz, 36)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "C", x="Pink abundance", y = "s(Pink abundance)")



# plot(pterm(herring.viz, 1)) +
#   l_ciBar(colour = "darkblue") + 
#   l_fitPoints(colour = "darkred") + 
#            l_rug(alpha = 0.3) +
#  l_fitLine(colour = "black") + 
#l_rug(mapping = aes(x=x, y=y), alpha = 0.8) +
#  l_ciLine(mul = 5, colour = "blue", linetype = 2) + 
#  l_points(shape = 19, size = 1, alpha = 0.1) + 

grid.sample <- gridPrint(herring.viz.term.1, herring.viz.term.2, herring.viz.term.3,
                         herring.viz.term.4, herring.viz.term.5, herring.viz.term.6, ncol=3, nrow=2)


ggsave("vars_gam_plot.png", grid.sample, device="png",width=10,height=7, dpi=350)

#test effect of te term
#te terms produce smooths of multiple predictors from tensor productos of any bases available for use with s
# The m= argument allows one to specify different types of covariance functions.


```



All species - reviewed, without multicolinear variables

```{r}

herring.basin.data.mod <- read_csv("herring_diet_basin_data_mod.csv") %>% 
 mutate(basin_id = as.factor(basin_id), atlantis_pred_group = as.factor(atlantis_pred_group),
        basin=as.factor(basin)) 

pred.year <- herring.basin.data.mod %>% 
  distinct(year,atlantis_pred_group) %>% 
  arrange(atlantis_pred_group,year)

pred.year %>% pull(year)

# examine data
# see https://cran.r-project.org/web/packages/fitdistrplus/vignettes/FAQ.html#can-i-personalize-the-default-plot-given-for-an-object-of-class-fitdist-or-fitdistcens

#gamma 1.4 to reduce overfitting

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

#full model with no interaction terms
base="tp"

herring.m1 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, bs=base, k = 12) +
                    s(atlantis_pred_group, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                   #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs = base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs = base, k=8) +
                    s(sockeye.ab, bs = base, k=8) +
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs = base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

saveRDS(herring.m1,"herring_gam_all_m1.rds")

gam.check(herring.m1)

concurvity(herring.m1, full = TRUE)

model.stats.all.m1 <- tidy(herring.m1) 


#model with interaction terms by year

base="tp"

herring.m2 <- gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(predator_size, by=year, bs=base, k = 10) + 
                    s(basin, by=year, bs='re', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(year, bs=base, k = 10) + 
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs = base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs = base, k=8) +
                    s(sockeye.ab, bs = base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs = base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


saveRDS(herring.m2,"herring_gam_all_m2.rds")

gam.check(herring.m2)

model.stats.all.m2 <- tidy(herring.m2) 



#model with interaction terms by basin

base="tp"

  
herring.m3 <- gam(proportion_sc ~
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(year, by=basin, bs=base, k=10) +
                    s(atlantis_pred_group, by=basin, bs='re', k=10) +
                    s(basin, bs='re', k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


concurvity(herring.m3, full = TRUE)

saveRDS(herring.m3,"herring_gam_all_m3.rds")

gam.check(herring.m3)

model.stats.all.m3 <- tidy(herring.m3) 

#combines year and basin
base="tp"

herring.m4 <- gam(proportion_sc ~ 
                   s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    s(year, by=basin, bs=base, k=10) +
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

#concurvity(herring.m4, full = TRUE)

saveRDS(herring.m4,"herring_gam_all_m4.rds")

gam.check(herring.m4)

model.stats.all.m4 <- tidy(herring.m4) 



#by predator group

base = "tp"

herring.m5 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(year, bs=base, k=10) +
                    s(predator_size, bs=base, k=10) + 
                    s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    #UPW +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


      
      #concurvity(herring.m4, full = TRUE)

saveRDS(herring.m5,"herring_gam_all_m5.rds")

gam.check(herring.m5)
      
model.stats.all.m5 <- tidy(herring.m5) 



#same as models 1-5 but uses ts as a basis

#full model with no interaction terms
base="ts"

herring.m6 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(predator_size, bs=base, k = 12) +
                    s(atlantis_pred_group, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                   #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs = base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs = base, k=8) +
                    s(sockeye.ab, bs = base, k=8) +
                    s(RR.SSTOut, bs = base, k=8) +
                    s(Chla.JDF, bs = base, k=8) +
                    s(FraserFlow.Out, bs = base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs = base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

saveRDS(herring.m6,"herring_gam_all_m6.rds")

gam.check(herring.m6)

concurvity(herring.m6, full = TRUE)

model.stats.all.m6 <- tidy(herring.m6) 


#model with interaction terms by year

base="ts"

herring.m7 <-  gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    #predator_size*year + 
                    s(predator_size, by=year, bs=base, k=10) +
                    s(basin, by=year, bs='re', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(year, bs=base, k=10) +
                    #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs = base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs = base, k=8) +
                    s(sockeye.ab, bs = base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs = base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


saveRDS(herring.m7,"herring_gam_all_m7.rds")

gam.check(herring.m7)
model.stats.all.m7 <- tidy(herring.m7) 



#model with interaction terms by basin

base="ts"

herring.m8 <- gam(proportion_sc ~ 
                    s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(year, by=basin, bs=base, k=10) +
                    s(atlantis_pred_group, by=basin, bs='re', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


concurvity(herring.m8, full = TRUE)

saveRDS(herring.m8,"herring_gam_all_m8.rds")

gam.check(herring.m8)

model.stats.all.m8 <- tidy(herring.m8) 


#combines year and basin
base="ts"

herring.m9 <- gam(proportion_sc ~ 
                   s(longitude, latitude, by=year, bs=base, k = 100) + 
                    s(atlantis_pred_group, by=basin, bs = 're', k=10) +
                    s(year, by=basin, bs=base, k=10) +
                    s(predator_size, by=basin, bs=base, k=10) + 
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(basin, bs='re', k=10) +
                    s(year, bs=base, k=10) +
                  #  s(predator_stage_id, bs = 'ts', k = 1) +
                   s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

#concurvity(herring.m9, full = TRUE)

saveRDS(herring.m9,"herring_gam_all_m9.rds")

gam.check(herring.m9)

model.stats.all.m9 <- tidy(herring.m9) 


#by predator group

base = "ts"

herring.m10 <- gam(proportion_sc ~ 
                    s(longitude, latitude, bs=base, k = 100) + 
                    s(year, bs=base, k=10) +
                    s(predator_size, bs=base, k=10) + 
                    s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                    s(atlantis_pred_group, by=year, bs = 're', k=10) +
                    s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


      
      #concurvity(herring.m10, full = TRUE)

saveRDS(herring.m10,"herring_gam_all_m10.rds")

gam.check(herring.m10)
      
model.stats.all.m10 <- tidy(herring.m10) 

base="ts"

herring.m11 <- gam(proportion_sc ~ 
                        s(longitude, latitude, by=year, bs=base, k = 100) + 
                        s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                        s(predator_size, by=year, bs=base, k = 10) +
                        s(year, by=basin, bs=base, k=10) +
                        s(atlantis_pred_group, by=year, bs = 're', k=10) +
                        s(year, by=atlantis_pred_group, bs = base, k=10) + 
                        basin +
                        year +
                     s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)


saveRDS(herring.m11,"herring_gam_all_m11.rds")

gam.check(herring.m11)
      
model.stats.all.m11 <- tidy(herring.m11) 


base="tp"

herring.m12 <- gam(proportion_sc ~ 
                        s(longitude, latitude, by=year, bs=base, k = 100) + 
                        s(basin, by=atlantis_pred_group, bs = 're', k=10) +
                        s(predator_size, by=year, bs=base, k = 10) +
                        s(year, by=basin, bs=base, k=10) +
                        s(atlantis_pred_group, by=year, bs = 're', k=10) +
                        s(year, by=atlantis_pred_group, bs = base, k=10) + 
                        basin +
                        year +
                        s(year, bs=base, k=10) +
                      s(other_herring, bs=base, k=8) +
                    s(SOG.Herring, bs=base, k=8) +
                    s(chum.ab, bs = base, k=8) +
                    s(chinook.ab, bs=base, k=8) +
                    s(coho.ab, bs=base, k=8) +
                    s(sockeye.ab, bs=base, k=8) +
                    s(RR.SSTOut, bs=base, k=8) +
                    s(Chla.JDF, bs=base, k=8) +
                    s(FraserFlow.Out, bs=base, k=8) +
                    s(PNI, bs=base, k=8) +
                    s(NPGO, bs=base, k=8) +
                    s(UPW, bs=base, k=8) +
                    s(MEI, bs=base, k=8),
                  family=betar(link="logit"), method="REML", gamma=1.4, select=TRUE, data=herring.basin.data.mod)

saveRDS(herring.m12,"herring_gam_all_m12.rds")

gam.check(herring.m12)
      
model.stats.all.m12 <- tidy(herring.m12) 


herring_gam_co_vars_m1 <- readRDS("~/herringdiet/herring_gam_all_m1.rds")
herring_gam_co_vars_m2 <- readRDS("~/herringdiet/herring_gam_all_m2.rds")
herring_gam_co_vars_m3 <- readRDS("~/herringdiet/herring_gam_all_m3.rds")
herring_gam_co_vars_m4 <- readRDS("~/herringdiet/herring_gam_all_m4.rds")
herring_gam_co_vars_m5 <- readRDS("~/herringdiet/herring_gam_all_m5.rds")
herring_gam_co_vars_m6 <- readRDS("~/herringdiet/herring_gam_all_m6.rds")
herring_gam_co_vars_m7 <- readRDS("~/herringdiet/herring_gam_all_m7.rds")
herring_gam_co_vars_m8 <- readRDS("~/herringdiet/herring_gam_all_m8.rds")
herring_gam_co_vars_m9 <- readRDS("~/herringdiet/herring_gam_all_m9.rds")
herring_gam_co_vars_m10 <- readRDS("~/herringdiet/herring_gam_all_m10.rds")
herring_gam_co_vars_m11 <- readRDS("~/herringdiet/herring_gam_all_m11.rds")
herring_gam_co_vars_m12 <- readRDS("~/herringdiet/herring_gam_all_m12.rds")

list.fits <- list(herring_gam_co_vars_m1,herring_gam_co_vars_m2,herring_gam_co_vars_m3,herring_gam_co_vars_m4,herring_gam_co_vars_m5,herring_gam_co_vars_m6,herring_gam_co_vars_m7,herring_gam_co_vars_m8, herring_gam_co_vars_m9, herring_gam_co_vars_m10, herring_gam_co_vars_m11, herring_gam_co_vars_m12)

model.names <- paste0("model_",1:length(list.fits))
no.fits <- 1:length(list.fits)

#aictab(cand.set = list.fits, modnames = model.names)

gam.res.table <- lapply(no.fits, gam_table, list.fits) %>% 
  bind_rows


write_csv(gam.res.table, "GAM_table_res_cov.csv")

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

#model 4

herring.m.final <- herring_gam_co_vars_m12

saveRDS(herring.m.final,"herring_gam_co_final.rds")



#https://cran.r-project.org/web/packages/mgcViz/vignettes/mgcviz.html
herring.viz <- getViz(herring.m.final)

print(plot(herring.viz, allTerms = T), pages = 1) # Calls print.plotGam()

#herring.m.draw <- draw(herring.m.final)+
  theme_classic()

herring.m.appraise <- appraise(herring.m.final)

#use to plot 2D
#plot(sm(herring.viz, 2)) + l_fitRaster() + l_fitContour() + l_points()

ggsave("co_gam_appraise.png", herring.m.appraise, device="png",width=10,height=8, dpi=350)


#use sm to extract smooth terms from model
herring.viz.term.1 <- plot(sm(herring.viz, 7)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "A", x="Gaussian quantiles", y = "Effects", title="s(basin):Predator group",
  subtitle="Dogfish")

herring.viz.term.2 <- plot(sm(herring.viz, 11)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
   labs(tag = "B", x="Gaussian quantiles", y = "Effects", title="s(basin):Predator group",
  subtitle="Large demersal fish")

herring.viz.term.3 <- plot(sm(herring.viz, 18)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "C", x="Gaussian quantiles", y = "Effects", title="s(basin):Predator group",
  subtitle="Orcas")

herring.viz.term.4 <- plot(sm(herring.viz, 25)) +
 l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "D", x="Predator size", y = "s(predator size): year")

herring.viz.term.5 <- plot(sm(herring.viz, 29)) +
  l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "E", x="Year", y = "s(year):basin", subtitle="San Juan Islands")

herring.viz.term.6 <- plot(sm(herring.viz, 40)) +
 l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "F", x="Year", y = "s(year):predator group", subtitle= "Dogfish")

herring.viz.term.7 <- plot(sm(herring.viz, 44)) +
 l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "G", x="Year", y = "s(year):predator group", subtitle= "Large demersal fish")

herring.viz.term.8 <- plot(sm(herring.viz, 51)) +
 l_ciPoly() +
  l_fitLine(colour = "black") + 
  theme_classic() +
  labs(tag = "H", x="Year", y = "s(year):predator group", subtitle= "Orca")


grid.sample <- gridPrint(herring.viz.term.1, herring.viz.term.2, herring.viz.term.3,  herring.viz.term.4, 
                         herring.viz.term.5, herring.viz.term.6, herring.viz.term.7, herring.viz.term.8, ncol=3, nrow=3)


ggsave("co_gam_plot.png", grid.sample, device="png",width=6,height=6, dpi=350)


#test effect of te term
#te terms produce smooths of multiple predictors from tensor productos of any bases available for use with s
# The m= argument allows one to specify different types of covariance functions.


```



